Realization of a Decoherence-Free Subspaces with Nitrogen-Vacancy Centers and $^{13}$C Nuclear Spins

### 2016-07-13 swap-nvst-paper/paper.tex
fixed missing label on a variable
fixed minor error
adjusted paper to focus on larmor control instead of a DFS SWAP

%%% math, symbols, etc.
\usepackage{physics,braket,bm,bbm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % square parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\hat{\v{#1}}} % unit vectors
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\Braket} % shorthand for braket notation

@@

\title{Noise-Protected Quantum Memory Registers with Nitrogen-Vacancy
  Centers and $^{13}$C Nuclear Spins}

@@

One such application of the NV center is as a memory register to store
quantum information. While typical control operations on the NV
electron spin happen on the scale of nanoseconds, its room-temperature
coherence time is on the scale of microseconds
\cite{dobrovitski2013quantum}. Dynamic decoupling schemes and the use
of isotopically pure diamond samples can increase the NV electron spin
coherence time by a few orders of magnitude
\cite{ryan2010robust}. Storing quantum information in the NV electron
spin itself, however, reserves a valuable resource, as the center
could otherwise be used to address nuclear spins and other quantum
processing resources. The weaker environmental coupling of $^{13}$C
nuclei present in diamond yields much longer coherence times than that
of the NV electron spin, and by entangling multiple nuclear spins into
the singlet-triplet states $\p{\ket{\u\d}\pm\ket{\d\u}}/\sqrt2$, it is
possible to generate logical qubits which are protected from various
sources of system and environmental dephasing noise. Such techniques
have been used to increase qubit lifetimes in both GaAs and silicon
quantum dots \cite{bluhm2011dephasing, maune2012coherent}, but have
yet to be realized in the NV center system.

In this paper, we propose using the singlet-triplet subspace of two
$^{13}$C nuclear spins with identical effective larmor frequencies as
a logical qubit for storing quantum memory. Here identical effective
larmor frequencies implies that, absent of symmetry-breaking control
fields, the nuclei are indistinguishable from the perspective of the
NV electron. This indistinguishability protects the logical qubit from
dephasing via coupling to the NV electron spin, potentially increasing
its coherence time by a few orders of magnitude. \fixme{elaborate} We
investigate the efficacy of this quantum memory register when realized
via the recently developed dynamical decoupling protocol dubbed
adaptive XY-$n$ (AXY-$n$) \cite{casanova2015robust}.

@@

When restricted to the subspace of $\set{\ket0,\ket{m_s}}$ with with
$m_s\in\set{1,-1}$ for the NV electron spin,
\begin{align}
  S_z = \f{m_s}2\p{I_\NV^z+\mathbbm1},
\end{align}
where $I_\NV^z\equiv\p{\op{m_s}-\op{0}}/2$ and
$\mathbbm1\equiv\op{m_s}+\op{0}$. Defining the interaction Hamiltonian
\begin{align}
  H_\t{int} = m_sI_\NV^z\sum_j\v A_j\c\v I_j
\end{align}
the effective larmor frequency
\begin{align}
  \v\omega_j = \gamma_j\v B - \f{m_s}2\v A_j,
\end{align}
and the effective Zeeman Hamiltonian
\begin{align}
  H_\t{nZ}^\t{eff} = -\sum_j\p{\gamma_j\v B - \f{m_s}2\v A_j}\c\v I_j
  = -\sum_j\v\omega_j\c\v I_j,
\end{align}
the system Hamiltonian can be written as
\begin{align}
  \tilde H_\t{sys} = H_\t{int} + H_\t{nZ}^\t{eff}.
  \label{H_sys_int_nZ}
\end{align}

Moving into the rotating frame of $H_\t{nZ}^\t{eff}$, the entire
system is described by the Hamiltonian
\begin{align}
  \tilde H_\t{int} = m_sI_\NV^z
  \sum_j\tilde{\v A_j}\p{t}\c\v I_j,
  \label{H_int_no_DD}
\end{align}
where $\tilde{\v A_j}\p{t}$ is the vector $\v A_j$ rotated about
$\uv\omega_j$ by an angle $\omega_jt$. Written explicitly,
\begin{align}
  \tilde{\v A_j}\p{t} = \v A_j^\parallel + \v A_j^\perp\cos\p{\omega_jt}
  + \v A_j^{\perp\perp}\sin\p{\omega_jt},
  \label{A_rot}
\end{align}
where
\begin{align}
  \v A_j^\parallel\equiv\p{\v A_j\c\uv\omega_j}\uv\omega_j
\end{align}
is the projection of $\v A_j$ along $\v \omega_j$,
\begin{align}
\v A_j^\perp\equiv\v A_j-\v A_j^\parallel
\end{align}
is the projection of $\v A_j$ onto the plane orthogonal to
$\v\omega_j$, and
\begin{align}
\v A_j^{\perp\perp}\equiv\uv\omega_j\times\v A_j^\perp
\end{align}
is the vector $\v A_j^\perp$ rotated about $\uv\omega_j$ by $\pi/2$.

@@

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $\sigma_z^\NV$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_\t{int} =
  F\p{t}m_sI_\NV^z\sum_j\tilde{\v A_j}\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_\t{int} = \f12 f_{k_{DD}} m_sI_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{H_int_resonance}
\end{align}
where $\uv n_j$ is a unit vector in the direction of $\v A_j^\perp$
rotated about $\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write (\ref{H_int_resonance}) with
\begin{align}
h_j^\t{int}\equiv f_{k_{DD}}m_sA_j^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{int} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\uv n_j.
  \label{H_int}
\end{align}

When the harmonics of $F\p{t}$ are off resonance with all $\omega_j$
and $\omega_{DD}\gg A_j$ for all $j$, the secular approximation allows
one to neglect $\tilde H_\t{int}$ entirely. In this case, one can
apply an auxiliary magnetic field
$\v B_\t{ctl}^{\p\omega}\p{t}=\v
B_\t{ctl}\cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}$ and shift into the
rotating frame of $H_\t{nZ}^\t{eff}$, much like the frame shift
between (\ref{H_sys_int_nZ}) and (\ref{H_int_no_DD}), to realize the
Hamiltonian
\begin{align}
  \tilde H_\t{sys}^\t{ctl} =
  -\sum_j\gamma_j\tilde{\v B}_{\t{ctl},j}\p{t}
  \cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}\c\v I_j,
\end{align}
with $\tilde{\v B}_{\t{ctl},j}\p{t}$ defined using $\v B_\t{ctl}$ and
$\v\omega_j$ the same way that $\tilde{\v A_j}\p{t}$ was defined using
$\v A_j$ and $\v\omega_j$ in (\ref{A_rot}), i.e. as $\v B_\t{ctl}$
rotated about $\uv\omega_j$ by an angle $\omega_jt$:
\begin{align}
  \tilde{\v B}_{\t{ctl},j}\p{t} = \v B_{\t{ctl},j}^\parallel
  + \v B_{\t{ctl},j}^\perp\cos\p{\omega_jt}
  + \v B_{\t{ctl},j}^{\perp\perp}\sin\p{\omega_jt},
\end{align}
where
\begin{align}
  \v B_{\t{ctl},j}^\parallel
  &\equiv \p{\v B_\t{ctl} \c\uv\omega_j}\uv\omega_j \\
  \v B_{\t{ctl},j}^\perp
  &\equiv \v B_\t{ctl} - \v B_{\t{ctl},j}^\parallel \\
  \v B_{\t{ctl},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_\t{ctl}^\perp.
\end{align}

When $\omega_\t{ctl}\gg\gamma_jB_\t{ctl}$, the secular approximation
allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and express
\begin{align}
  \tilde H_\t{ctl} = -\f12\sum_{j:\omega_j=\omega_\t{ctl}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_\t{ctl}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_\t{ctl} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_\t{ctl}.
\end{align}
One can concisely write (\ref{H_ctl_resonance}) with
\begin{align}
h_j^\t{ctl}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{ctl} = \sum_{j:\omega_j=\omega_\t{ctl}}
  h_j^\t{ctl}\v I_j\c\uv m_j.
  \label{H_ctl}
\end{align}

In both (\ref{H_int}) and (\ref{H_ctl}), one can choose $\phi_{DD}$
and $\phi_\t{ctl}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_\t{ctl}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^\t{int}$ and $h_j^\t{ctl}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_\t{ctl}$. These controls, together with rotations of the NV
electron spin, allow one to realize a universal set of quantum
gates. An exception to this statement is when two or more nuclei have
the same effective larmor frequency $\omega_j$, in which case they are
addressed simultaneously in (\ref{H_int}) and (\ref{H_ctl}).

@@

We call two spins indexed by $j$ and $k$ elements of a \emph{larmor
  set} if their effective larmor frequencies are equal,
i.e. $\omega_j=\omega_k$, and refer to a larmor set with exactly two
spins as a \emph{larmor pair}. The problem with individually
addressing elements of a larmor set is manifest in (\ref{H_int}) and
(\ref{H_ctl}), wherein the previously developed procedure for nuclear
spin addressing does not single out any nuclei. Nonetheless, under
certain conditions it is possible to individually address elements of
a larmor set by applying additional control fields.

@@

In the frame of $H_\t{nZ}^\t{eff}=-\sum_j\v\omega_j\c\v I_j$, the
decoupling Hamiltonian becomes
\begin{align}
  \tilde H_\t{dec} = -\sum_j\gamma_j\tilde{\v B}_{\t{dec},j}
  \c\v I_j \cos\p{\omega_rt-\phi_\t{dec}}
  \label{H_dec_eff_full}
\end{align}
with
\begin{align}
  \tilde{\v B}_{\t{dec},j} = \v B_{\t{dec},j}^\parallel
  + \v B_{\t{dec},j}^\perp\cos\p{\omega_j t}
  + \v B_{\t{dec},j}^{\perp\perp}\sin\p{\omega_j t}
\end{align}
and
\begin{align}
  \v B_{\t{dec},j}^\parallel
  &\equiv \p{\v B_\t{dec} \c\uv\omega_j}\uv\omega_j, \\
  \v B_{\t{dec},j}^\perp
  &\equiv \v B_\t{dec} - \v B_{\t{dec},j}^\parallel, \\
  \v B_{\t{dec},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_\t{dec}^\perp.
\end{align}
If $\omega_r,\omega_j\gg \gamma_jB_\t{dec}$, then we can neglect
$\v B_{\t{dec},j}^\parallel$ in (\ref{H_dec_eff_full}) by the secular
approximation, and by the rotating approximation express
\begin{multline}
  \tilde H_\t{dec} \\
  = -\sum_{j:\omega_j=\omega_r}
  \f12\gamma_j\p{\v B_{\t{dec},j}^\perp\cos\phi_\t{dec}
  + \v B_{\t{dec},j}^{\perp\perp}\sin\phi_\t{dec}}\c\v I_j.
\end{multline}
Defining
\begin{align}
  \v\nu_j = \f12\gamma_j\p{\v B_{\t{dec},j}^\perp\cos\phi_\t{dec}
  + \v B_{\t{dec},j}^{\perp\perp}\sin\phi_\t{dec}},
\end{align}
we can express
\begin{align}
  \tilde H_\t{dec} = -\sum_{j:\omega_j=\omega_r}\v\nu_j\c\v I_j.
\end{align}
The system Hamiltonian in the frame of $H_\t{nZ}^\t{eff}$ thus takes
the form
\begin{align}
  \tilde H_\t{sys}^\t{dec} = \tilde H_\t{int} + \tilde H_\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  \p{h_j^\t{int}I_\NV^z\v I_j\c\uv n_j - \v\nu_j\c\v I_j}.
\end{align}

In the frame of $\tilde H_\t{dec}$, the entire system is governed by
the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\tilde{\v n}_j\p{t},
  \label{H_int_dec_full}
\end{align}
where $\tilde{\v n}_j\p{t}$ is the unit vector $\uv n_j$ rotated
about $\uv\nu_j$ by $\nu_jt$, i.e.
\begin{align}
  \tilde{\v n}_j\p{t} = \v n_j^\parallel +
  \v n_j^\perp\cos\p{\nu_jt} + \v n_j^{\perp\perp}\sin\p{\nu_jt},
\end{align}
with
\begin{align}
  \v n_j^\parallel &\equiv \p{\uv n_j\c\uv\nu_j}\uv\nu_j, \\
  \v n_j^\perp &\equiv \uv n_j - \uv n_j^\parallel, \\
  \v n_j^{\perp\perp} &\equiv \uv\nu_j\times\uv n_j.
\end{align}
If $\omega_j\gg\nu_j\gg h_j^\t{int}$, the secular approximation allows
us to neglect $\v n_j^\perp$ and $\v n_j^{\perp\perp}$ in
(\ref{H_int_dec_full}), leaving
\begin{align}
  \tilde H_\t{int}^\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\v n_j^\parallel
  = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}I_\NV^z\v I_j\c\uv\nu_j,
  \label{H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int} = h_j\uv n_j\c\uv\nu_j.
\end{align}

>>>>

%%% math, symbols, etc.
\usepackage{physics,braket,bm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % square parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\hat{\v{#1}}} % unit vectors
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\Braket} % shorthand for braket notation

@@

\title{Realization of a Decoherence-Free Subspaces with
  Nitrogen-Vacancy Centers and $^{13}$C Nuclear Spins}

@@

One such application of the NV center is as a memory register to store
quantum information. While typical control operations on the NV
electron spin happen on the scale of nanoseconds, its room-temperature
coherence time is on the scale of microseconds
\cite{dobrovitski2013quantum}. Dynamic decoupling schemes and the use
of isotopically pure diamond samples can increase the NV electron spin
coherence time by a few orders of magnitude
\cite{ryan2010robust}. Storing quantum information in the NV electron
spin itself, however, reserves a valuable resource, as the center
could otherwise be used to address nuclear spins and other quantum
processing resources. The weaker environmental coupling of $^{13}$C
nuclei present in diamond yields much longer coherence times than that
of the NV electron spin, and by entangling multiple nuclear spins,
e.g. into the singlet-triplet states
$\p{\ket{\u\d}\pm\ket{\d\u}}/\sqrt2$, it is possible to generate
logical qubits which are protected from various sources of system and
environmental dephasing noise.

In this letter, we present a means to address individual $^{13}$C
nuclear spins which are normally indistinguishable due to symmetric
coupling to the NV center. We use this addressing scheme to initialize
qubits in a decoherence-free subspace (DFS) of nuclear spins, and
propose a means for realizing a coherent SWAP operation between the NV
center electron spin state and the DFS qubit.

\section{Addressing symmetric $^{13}$C nuclear spins}

Restricting the NV electron spin states to the subspace spanned by
$\set{\ket0,\ket{m_s}}$ ($m_s=\pm1$), the AXY-$n$ pulse sequence
developed in by Casanova et al. in \cite{casanova2015robust} and
control scheme developed by Wang et al. in \cite{wang2015positioning}
allows addressing individual $^{13}$C nuclear spins via Hamiltonians
of the form
\begin{align}
  H_\t{int} = h_\t{int} I_\NV^z\v I_j\c\uv n_j,
  \label{H_int}
\end{align}
for a tunable coupling strength $h_\t{int}$ and direction $\uv n_j$,
where $I_\NV^z\equiv\p{\op{m_s}-\op{0}}/2$ and $\v I_j$ is the spin
operator for nuclear spin $j$. The protocol for realizing this
effective Hamiltonian, however, is conditional on a sufficiently
unique larmor frequency $\omega_j$, i.e. that for all $n\ne j$
\begin{align}
  \abs{\omega_n-\omega_j} \gg \abs{f_{k_{DD}}\v A_j^\perp},
\end{align}
where $f_{k_{DD}}$ is the fourier component of the AXY-$n$ protocol on
resonance with $\omega_j$ and $\v A_j^\perp$ is the component of the
hyperfine vector $\v A_j$ perpendicular to the effective larmor
precession axis $\v\omega_j$.

It is possible, however, for two nuclear spins (indexed by 1 and 2) to
have identical larmor frequencies $\omega_1=\omega_2\equiv\omega$,
which occurs when they couple symmetrically to the NV center with
$A_1=A_2$. The state space of two such spins, which we call {\it
  larmor pairs}, contains a decoherence-free subspace (DFS) with
respect to spin-spin coupling to the NV center. This DFS is spanned by
the singlet-triplet states $\p{\ket{\u\d}\pm\ket{\d\u}}/\sqrt2$, or
equivalently by $\set{\ket{\u\d},\ket{\d\u}}$.

When trying to address one spin in a larmor pair, the protocols for
(\ref{H_int}) will simultaneously couple $I_\NV^z$ to both
$\v I_1\c\uv n_1$ and $\v I_2\c\uv n_2$ with identical coupling
strength ($h_\t{int}$). So long as $\uv n_1$ and $\uv n_2$
(equivalently, $\v A_1^\perp$ and $\v A_2^\perp$) are not parallel,
however, it is possible to address only one of these spins at a time
by applying additional control fields. In particular, the application
of a magnetic decoupling field
$\v B_\t{dec}^{\p\omega}\p{t}=\v B_\t{dec}\cos\p{\omega
  t-\phi_\t{dec}}$ during the AXY-$n$ addressing protocol suppresses
components of $\uv n_j$ which are perpendicular to the vector
$\v B_\t{dec}$ projected onto the plane perpendicular to precession
axis $\v\omega_j$ and rotated about this axis by $\phi_\t{dec}$. One
can therefore choose a direction $\v B_\t{dec}$ which completely
suppresses NV electron spin coupling to spin 1 (2), leaving only
coupling to spin 2 (1).

\fixme{Comment on how likely are we to find larmor pairs?}

\section{Qubit initialization in a decoherence-free subspace}

\bibliography{\jobname}

\pagebreak
\clearpage
\widetext
\begin{center}
  \textbf{\large Supplementary Information for ``Delayed entanglement
    echo for individual control of a large number of nuclear spins''}
\end{center}
\setcounter{equation}{0} \setcounter{figure}{0} \setcounter{table}{0}
\setcounter{page}{1} \makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibnumfmt}[1]{[S#1]}
\renewcommand{\citenumfont}[1]{S#1}

\tableofcontents{}

@@

When restricted to the subspace of $\set{\ket0,\ket{m_s}}$ with with
$m_s\in\set{1,-1}$ for the NV electron spin,
\begin{align}
  S_z = \f{m_s}2\p{I_\NV^z+\mathbf1},
\end{align}
where $I_\NV^z\equiv\p{\op{m_s}-\op{0}}/2$ and
$\mathbf1\equiv\op{m_s}+\op{0}$. Defining the interaction Hamiltonian
\begin{align}
  H_\t{int} = m_sI_\NV^z\sum_j\v A_j\c\v I_j
\end{align}
the effective larmor frequency
\begin{align}
  \v\omega_j = \gamma_j\v B - \f{m_s}2\v A_j,
\end{align}
and the effective Zeeman Hamiltonian
\begin{align}
  H_\t{nZ}^\t{eff} = -\sum_j\p{\gamma_j\v B - \f{m_s}2\v A_j}\c\v I_j
  = -\sum_j\v\omega_j\c\v I_j,
\end{align}
the system Hamiltonian can be written as
\begin{align}
  \tilde H_\t{sys} = H_\t{int} + H_\t{nZ}^\t{eff}.
  \label{S_H_sys_int_nZ}
\end{align}

Moving into the rotating frame of $H_\t{nZ}^\t{eff}$, the entire
system is described by the Hamiltonian
\begin{align}
  \tilde H_\t{int} = m_sI_\NV^z
  \sum_j\tilde{\v A_j}\p{t}\c\v I_j,
  \label{S_H_int_no_DD}
\end{align}
where $\tilde{\v A_j}\p{t}$ is the vector $\v A_j$ rotated about
$\uv\omega_j$ by an angle $\omega_jt$. Written explicitly,
\begin{align}
  \tilde{\v A_j}\p{t} = \v A_j^\parallel + \v A_j^\perp\cos\p{\omega_jt}
  + \v A_j^{\perp\perp}\sin\p{\omega_jt},
  \label{S_A_rot}
\end{align}
where
\begin{align}
  \v A_j^\parallel\equiv\p{\v A_j\c\uv\omega_j}\uv\omega_j
\end{align}
is the projection of $\v A_j$ along $\v \omega_j$,
\begin{align}
  \v A_j^\perp\equiv\v A_j-\v A_j^\parallel
\end{align}
is the projection of $\v A_j$ onto the plane orthogonal to
$\v\omega_j$, and
\begin{align}
  \v A_j^{\perp\perp}\equiv\uv\omega_j\times\v A_j^\perp
\end{align}
is the vector $\v A_j^\perp$ rotated about $\uv\omega_j$ by $\pi/2$.

@@

Dynamic decoupling schemes effectively prepend a modulation function
\begin{align}
  F\p{t}=\sum_kf_k\cos\p{k\omega_{DD}t-\phi_{DD}}
\end{align}
with fundamental frequency $\omega_{DD}$, phase $\phi_{DD}$, and range
$\set{1,-1}$ to $I_\NV^z$, so that the system Hamiltonian reads
\begin{align}
  \tilde H_\t{int} =
  F\p{t}m_sI_\NV^z\sum_j\tilde{\v A_j}\p{t}\c\v I_j.
\end{align}
At a resonance $k_{DD}\omega_{DD}\equiv\omega_r\gg A_j$, the secular
approximation allows one to neglect $\v A_j^\parallel$ and express
\begin{align}
  \tilde H_\t{int} = \f12 f_{k_{DD}} m_sI_\NV^z
  \sum_{j:\omega_j=\omega_r} A_j^\perp\v I_j\c\uv n_j,
  \label{S_H_int_resonance}
\end{align}
where $\uv n_j$ is a unit vector in the direction of $\v A_j^\perp$
rotated about $\uv\omega_j$ by $\phi_{DD}$, i.e.
\begin{align}
  \uv n_j = \uv A_j^\perp\cos\phi_{DD} +
  \uv A_j^{\perp\perp}\sin\phi_{DD}.
\end{align}
One can concisely write (\ref{S_H_int_resonance}) with
\begin{align}
  h_j^\t{int}\equiv f_{k_{DD}}m_sA_j^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{int} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\uv n_j.
  \label{S_H_int}
\end{align}

When the harmonics of $F\p{t}$ are off resonance with all $\omega_j$
and $\omega_{DD}\gg A_j$ for all $j$, the secular approximation allows
one to neglect $\tilde H_\t{int}$ entirely. In this case, one can
apply an auxiliary magnetic field
$\v B_\t{ctl}^{\p\omega}\p{t}=\v
B_\t{ctl}\cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}$ and shift into the
rotating frame of $H_\t{nZ}^\t{eff}$, much like the frame shift
between (\ref{S_H_sys_int_nZ}) and (\ref{S_H_int_no_DD}), to realize
the Hamiltonian
\begin{align}
  \tilde H_\t{sys}^\t{ctl} =
  -\sum_j\gamma_j\tilde{\v B}_{\t{ctl},j}\p{t}
  \cos\p{\omega_\t{ctl}t-\phi_\t{ctl}}\c\v I_j,
\end{align}
with $\tilde{\v B}_{\t{ctl},j}\p{t}$ defined using $\v B_\t{ctl}$ and
$\v\omega_j$ the same way that $\tilde{\v A_j}\p{t}$ was defined using
$\v A_j$ and $\v\omega_j$ in (\ref{S_A_rot}), i.e. as $\v B_\t{ctl}$
rotated about $\uv\omega_j$ by an angle $\omega_jt$:
\begin{align}
  \tilde{\v B}_{\t{ctl},j}\p{t} = \v B_{\t{ctl},j}^\parallel
  + \v B_{\t{ctl},j}^\perp\cos\p{\omega_jt}
  + \v B_{\t{ctl},j}^{\perp\perp}\sin\p{\omega_jt},
\end{align}
where
\begin{align}
  \v B_{\t{ctl},j}^\parallel
  &\equiv \p{\v B_\t{ctl} \c\uv\omega_j}\uv\omega_j \\
  \v B_{\t{ctl},j}^\perp
  &\equiv \v B_\t{ctl} - \v B_{\t{ctl},j}^\parallel \\
  \v B_{\t{ctl},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_\t{ctl}^\perp.
\end{align}

When $\omega_\t{ctl}\gg\gamma_jB_\t{ctl}$, the secular approximation
allows one to neglect $\v B_{\t{ctl},j}^\parallel$ and express
\begin{align}
  \tilde H_\t{ctl} = -\f12\sum_{j:\omega_j=\omega_\t{ctl}}
  \gamma_jB_{\t{ctl},j}^\perp \v I_j\c\uv m_j,
  \label{S_H_ctl_resonance}
\end{align}
where $\uv m_j$ is the direction of $\v B_{\t{ctl},j}^\perp$ rotated
about $\uv\omega_j$ by $\phi_\t{ctl}$, i.e.
\begin{align}
  \uv m_j = \uv B_{\t{ctl},j}^\perp\cos\phi_\t{ctl} +
  \uv B_{\t{ctl},j}^{\perp\perp}\sin\phi_\t{ctl}.
\end{align}
One can concisely write (\ref{S_H_ctl_resonance}) with
\begin{align}
  h_j^\t{ctl}\equiv-\gamma_jB_{\t{ctl},j}^\perp/2
\end{align}
as
\begin{align}
  \tilde H_\t{ctl} = \sum_{j:\omega_j=\omega_\t{ctl}}
  h_j^\t{ctl}\v I_j\c\uv m_j.
  \label{S_H_ctl}
\end{align}

In both (\ref{S_H_int}) and (\ref{S_H_ctl}), one can choose
$\phi_{DD}$ and $\phi_\t{ctl}$ such that $\uv n_j\p\phi$ and
$\uv m_j\p{\phi_\t{ctl}}$ lie anywhere in the plane perpendicular to
$\uv\omega_j$. One can also tune $h_j^\t{int}$ and $h_j^\t{ctl}$
respectively via free parameters in the AXY-$n$ sequence and
$\v B_\t{ctl}$. These controls, together with rotations of the NV
electron spin, allow one to realize a universal set of quantum
gates. An exception to this statement is when two or more nuclei have
the same effective larmor frequency $\omega_j$, in which case they are
addressed simultaneously in (\ref{S_H_int}) and (\ref{S_H_ctl}).

@@

We call two spins indexed by $j$ and $k$ elements of a \emph{larmor
  set} if their effective larmor frequencies are equal,
i.e. $\omega_j=\omega_k$, and refer to a larmor set with exactly two
spins as a \emph{larmor pair}. The problem with individually
addressing elements of a larmor set is manifest in (\ref{S_H_int}) and
(\ref{S_H_ctl}), wherein the previously developed procedure for
nuclear spin addressing does not single out any nuclei. Nonetheless,
under certain conditions it is possible to individually address
elements of a larmor set by applying additional control fields.

@@

In the frame of $H_\t{nZ}^\t{eff}=-\sum_j\v\omega_j\c\v I_j$, the
decoupling Hamiltonian becomes
\begin{align}
  \tilde H_\t{dec} = -\sum_j\gamma_j\tilde{\v B}_{\t{dec},j}
  \c\v I_j \cos\p{\omega_rt-\phi_\t{dec}}
  \label{S_H_dec_eff_full}
\end{align}
with
\begin{align}
  \tilde{\v B}_{\t{dec},j} = \v B_{\t{dec},j}^\parallel
  + \v B_{\t{dec},j}^\perp\cos\p{\omega_j t}
  + \v B_{\t{dec},j}^{\perp\perp}\sin\p{\omega_j t}
\end{align}
and
\begin{align}
  \v B_{\t{dec},j}^\parallel
  &\equiv \p{\v B_\t{dec} \c\uv\omega_j}\uv\omega_j, \\
  \v B_{\t{dec},j}^\perp
  &\equiv \v B_\t{dec} - \v B_{\t{dec},j}^\parallel, \\
  \v B_{\t{dec},j}^{\perp\perp}
  &\equiv \uv\omega_j\times\v B_\t{dec}^\perp.
\end{align}
If $\omega_r,\omega_j\gg \gamma_jB_\t{dec}$, then we can neglect
$\v B_{\t{dec},j}^\parallel$ in (\ref{S_H_dec_eff_full}) by the
secular approximation, and by the rotating approximation express
\begin{align}
  \tilde H_\t{dec} = -\sum_{j:\omega_j=\omega_r}
  \f12\gamma_j\p{\v B_{\t{dec},j}^\perp\cos\phi_\t{dec} +
  \v B_{\t{dec},j}^{\perp\perp}\sin\phi_\t{dec}}\c\v I_j.
\end{align}
Defining
\begin{align}
  \v\nu_j = \f12\gamma_j\p{\v B_{\t{dec},j}^\perp\cos\phi_\t{dec}
  + \v B_{\t{dec},j}^{\perp\perp}\sin\phi_\t{dec}},
\end{align}
we can express
\begin{align}
  \tilde H_\t{dec} = -\sum_{j:\omega_j=\omega_r}\v\nu_j\c\v I_j.
\end{align}
The system Hamiltonian in the frame of $H_\t{nZ}^\t{eff}$ thus takes
the form
\begin{align}
  \tilde H_\t{sys}^\t{dec} = \tilde H_\t{int} + \tilde H_\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  \p{h_j^\t{int}I_\NV^z\v I_j\c\uv n_j - \v\nu_j\c\v I_j}.
\end{align}

In the frame of $\tilde H_\t{dec}$, the entire system is governed by
the Hamiltonian
\begin{align}
  \tilde H_\t{int}^\t{dec} = \sum_{j:\omega_j=\omega_r}
  h_j^\t{int}I_\NV^z\v I_j\c\tilde{\v n}_j\p{t},
  \label{S_H_int_dec_full}
\end{align}
where $\tilde{\v n}_j\p{t}$ is the unit vector $\uv n_j$ rotated about
$\uv\nu_j$ by $\nu_jt$, i.e.
\begin{align}
  \tilde{\v n}_j\p{t} = \v n_j^\parallel +
  \v n_j^\perp\cos\p{\nu_jt} + \v n_j^{\perp\perp}\sin\p{\nu_jt},
\end{align}
with
\begin{align}
  \v n_j^\parallel &\equiv \p{\uv n_j\c\uv\nu_j}\uv\nu_j, \\
  \v n_j^\perp &\equiv \uv n_j - \uv n_j^\parallel, \\
  \v n_j^{\perp\perp} &\equiv \uv\nu_j\times\uv n_j.
\end{align}
If $\omega_j\gg\nu_j\gg h_j^\t{int}$, the secular approximation allows
us to neglect $\v n_j^\perp$ and $\v n_j^{\perp\perp}$ in
(\ref{S_H_int_dec_full}), leaving
\begin{align}
  \tilde H_\t{int}^\t{dec}
  = \sum_{j:\omega_j=\omega_r}
  \tilde h_j^\t{int}I_\NV^z\v I_j\c\uv\nu_j,
  \label{S_H_int_dec}
\end{align}
where
\begin{align}
  \tilde h_j^\t{int} = h_j^\t{int}\uv n_j\c\uv\nu_j.
\end{align}

In words, the decoupling field $\v B_\t{dec}^{\p\omega}\p{t}$ modifies
$\tilde H_\t{int}$ in (\ref{S_H_int}) by suppressing the components of
$\uv n_j$ (i.e. $\uv A_j^\perp$ rotated about $\uv\omega_j$ by
$\phi_{DD}$) which are perpendicular to the respective $\uv\nu_j$
(i.e.  $\uv B_{\t{dec},j}^\perp$ rotated about $\uv\omega_j$ by
$\phi_\t{dec}$). The utility of this decoupling field lies in the fact
that when $\uv n_j$ and $\uv\nu_j$ are orthogonal, the interaction
with spin $j$ is suppressed as $\tilde h_j^\t{int}=0$. Considering a
single larmor pair, one can therefore apply a decoupling field to
eliminate one of the two terms in (\ref{S_H_int_dec}) and realize an
effective Hamiltonian which couples the NV electron spin to only one
nuclear spin. This scheme fails for larmor pairs with mutually
parallel $\v A_j^\perp$.

The tools developed in this paper do not allow for single-spin
operations on elements of a larmor pair. Such operations must
therefore be either forgone, or else realized via composite operations
making use of (\ref{S_H_ctl}) and (\ref{S_H_int_dec}). The same is
true of spin-spin coupling between the NV electron spin and a single
element of a larmor set with more than two nuclei.
